---
title: "Stat301_Project"
author: "Sasha Johnson"
date: "2025-04-18"S
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Dataset
```{r Import Dataset}
library(readxl)
Homework3_stat301 <- read_excel("~/Documents/Metro_State/Spring_2025/Stat301/Stat301_project/Homework3_stat301.xlsx")
Homework3_stat301
```

Let SP = Scoring Points
$H_0: \mu_\text{SPcenter} = \mu_\text{SPlw} = \mu_\text{SPrw} = \mu_\text{SPd}$ 
vs 
$H_a: \text{At least 2 mean scoring points differ}$
Perform One-Way ANOVA
```{r Perform One-Way ANOVA}
# PTS stands for scoring points
scoring_points_model = aov(PTS~Pos, data = Homework3_stat301)
summary(scoring_points_model)
```
p-value = <2e-16 (i.e. 0.0000000000000002); f-value = 68.39
Reject H0; there is no evidence to support the claim that the mean scoring 
points are equal.
Reject H0; there is evidence to support the claim that at least 2 scoring point 
means differ.

Because our P-value was less then 0.05 we can reject the null hypotheses and accept the alternative hypothesis.  At least two means (positions) differ in point production
