---
title: "Stat301_Project"
author: "Sasha Johnson"
date: "2025-04-18"S
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Dataset
```{r Import Dataset}
library(readxl)
player_stats <- read_excel("~/Documents/Metro_State/Spring_2025/Stat301/Stat301_project/player_stats.xlsx")
# View(player_stats)
```

Clean Dataset
```{r Clean Dataset}
library(tidyverse)

# Truncates player positions if they have multiple positions
# ex: "LW/RW" turns into "LW"
player_stats$Pos = sub("/.*", "", player_stats$Pos)

# Keep only applicable positions (LW, C, RW, D)
#Remove any players that have less than 30 games played - for both questions
player_stats = player_stats %>%
    filter(GP > 30) %>%
    filter(Pos %in% c("LW", "C", "RW", "D"))
```

$H_0: \mu_\text{CenterPTS} = \mu_\text{LeftWingPTS} = \mu_\text{RightWingPTS} = \mu_\text{DefensemanPTS}$ 
vs 
$H_a: \text{At least 2 mean scoring points differ}$
Perform One-Way ANOVA
```{r Perform One-Way ANOVA}
library(ggplot2)
library(tidyverse)
# Select single season for one-way anova
player_stats_oneway = player_stats %>%
    filter(Season == 2016)

# PTS stands for scoring points
scoring_points_model = aov(PTS~Pos, data = player_stats_oneway)
summary(scoring_points_model)

# Create point graph with error bars
player_stats_oneway %>%
    group_by(Pos) %>%
    summarise(mean_pts = mean(PTS),
              se_pts = sd(PTS)/sqrt(length(PTS)),
              t_star = qt(1 - 0.05/2, length(PTS) - 1)
              ) %>%
    ggplot(aes(x = Pos, y = mean_pts)) +
    geom_point() +
    geom_errorbar(aes(ymin = mean_pts - se_pts * t_star,
                      ymax = mean_pts + se_pts * t_star)
                  ) +
    scale_y_continuous(limits = c(0, NA)) +
    labs(x = "Player Position",
         y = "Average Points",
         title = "Points by Position for the 2016 season")
```
p-value = <1.15e-09; f-value = 15.4
Reject H0; there is evidence to support the claim that at least 2 scoring point means differ.


### Second Question
Two-Way ANOVA
$H_0: \text{There is no interaction between player age and player position}$
$H_0: \mu_\text{CenterPTS} = \mu_\text{LeftWingPTS} = \mu_\text{RightWingPTS} = \mu_\text{DefensmanPTS}$
$H_0: \mu_\text{Young} = \mu_\text{Prime} = \mu_\text{Old}$
vs 
$H_a: \text{There is an interaction between player age and player position}$
$H_a: \text{At least mean differs for player age}$
$H_a: \text{At least mean differs for player position}$

Perform Two-Way ANOVA
```{r Perform Two-Way ANOVA}
library(tidyverse)
library(ggplot2)
# Add column to categorize age into 3 categories
player_stats$AgeGroup = cut(player_stats$Age,
                            breaks = c(19, 25, 31, 37),
                            labels = c("young", "prime", "old"),
                            right = TRUE)

# We'll have N/A in some AgeGroups - removing them
player_stats_twoway = subset(player_stats, !is.na(AgeGroup)) %>%
    filter(Season == 2017)
# View(player_stats_twoway)

age_position_model = aov(PTS ~ Pos + AgeGroup + Pos * AgeGroup, data = player_stats_twoway)
summary(age_position_model)
#TukeyHSD(age_position_model)

# graphing
player_stats_twoway %>%
    group_by(Pos, AgeGroup) %>%
    summarise(mean_pts = mean(PTS),
              se_pts = sd(PTS)/sqrt(length(PTS)),
              t_star = qt(1 - 0.05/2, length(PTS) - 1),
              .groups = "drop"
              ) %>%
    ggplot(aes(x = Pos,
               y = mean_pts,
               fill = AgeGroup)) +
    geom_bar(stat = "identity",
             position = "dodge") +
    geom_errorbar(aes(ymin = mean_pts - se_pts * t_star,
                      ymax = mean_pts + se_pts * t_star),
                  position = "dodge") +
    # I put the text here because I thought it looked incorrect - remove for the paper
    geom_text(
        aes(label = round(mean_pts, 1)),
        position = position_dodge(0.9)
        ) +
    labs(x = "Position",
         y = "Average Points",
         title = "Points by Position and Age Group for the 2017 season",
         fill = "Age Group")

```
              Df Sum Sq Mean Sq F value   Pr(>F)    
Pos            3  16801    5600  16.489 2.71e-10 ***
AgeGroup       2   1128     564   1.661    0.191    
Pos:AgeGroup   6   1219     203   0.598    0.732    
Residuals    579 196654     340      

Position:
F-value: 16.489; P-value: 2.71e-10
Because our p-value is less than 0.05 we reject the null hypothesis and accept the alternative hypothesis.

AgeGroup
F-value: 1.661; P-value: 0.191
Because our P-value is greater than 0.05 we fail to reject the null hypothesis

Pos : AgeGroup
F-value: 0.598; P-value: 0.732
Because our P-value is greater than 0.05 we fail to reject the null hypothesis